---
output: github_document
keep_md: TRUE
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

[![CRAN](http://www.r-pkg.org/badges/version/R2BEAT)](https://cran.r-project.org/package=R2BEAT)
[![Downloads](http://cranlogs.r-pkg.org/badges/R2BEAT?color=brightgreen)](http://www.r-pkg.org/pkg/R2BEAT)
[![Mentioned in Awesome Official Statistics](https://awesome.re/mentioned-badge.svg)](http://www.awesomeofficialstatistics.org)

# R2BEAT: Multistage Sampling Design and Sample Selection

Multivariate optimal allocation for different domains in one and two stages stratified sample design. R2BEAT extends the Neyman (1934) – Tschuprow (1923) allocation method to the case of several variables, adopting a generalization of the Bethel’s proposal (1989). R2BEAT develops this methodology but, moreover, it allows to determine the sample allocation in the multivariate and multi-domains case of estimates for two-stage stratified samples. It also allows to perform both Primary Stage Units and Secondary Stage Units selection.

*R2BEAT* easily manages all the complexity due to the optimal sample allocation in two-stage sampling design and provides several outputs for evaluating the allocation.
Its name stands for "R 'to' Bethel Extended Allocation for Two-stage". 
It is an extension of another open-source software called *Mauss-R* (Multivariate Allocation of Units in Sampling Surveys), implemented by ISTAT  researchers (https://www.istat.it/en/methods-and-tools/methods-and-it-tools/design/design-tools/mauss-r). *Mauss-R* determines the optimal sample allocation in multivariate and multi-domains estimation, for one-stage stratified samples. 

To complete the suite of tools developed by Istat in order to cover the stratified sample design, we cite *SamplingStrata* (https://CRAN.R-project.org/package=SamplingStrata), that allows to jointly optimize both the stratification of the sampling frame and the allocation, still in the multivariate multidomain case (only for one-stage designs), and *MultiWay.Sample.Allocation*, that allows to determine the optimal sample allocation for multi-way stratified sampling designs and incomplete stratified sampling designs (https://www.istat.it/en/methods-and-tools/methods-and-it-tools/design/design-tools/multiwaysampleallocation).

For a complete illustration of the methodology see the vignette "R2BEAT methodology and use" (https://barcaroli.github.io/R2BEAT/articles/R2BEAT_methodology.html). 

A complete example is illustrated in the vignette "Two-stage sampling design workflow" (https://barcaroli.github.io/R2BEAT/articles/R2BEAT_workflow.html). 


## Installation

You can install the released version of R2BEAT from [CRAN](https://CRAN.R-project.org) with:

``` {r, eval=FALSE, echo=TRUE}
install.packages("R2BEAT")
```

or the last version of R2BEAT from github with:
``` {r, eval=FALSE, echo=TRUE}
install.packages("devtools")
devtools::install_github("barcaroli/R2BEAT")
```

## Example

``` {r, eval=FALSE, echo=TRUE}
library(R2BEAT)

#-------------------------------------------------------------------------------
# Read sampling frame
library(readr)
pop <- read_rds("https://github.com/barcaroli/R2BEAT_workflows/blob/master/pop.RDS?raw=true")
str(pop)
# 'data.frame':	2258507 obs. of  13 variables:
#   $ region       : Factor w/ 3 levels "north","center",..: 1 1 1 1 1 1 1 1 1 1 ...
# $ province     : Factor w/ 6 levels "north_1","north_2",..: 1 1 1 1 1 1 1 1 1 1 ...
# $ municipality : num  1 1 1 1 1 1 1 1 1 1 ...
# $ id_hh        : Factor w/ 963018 levels "H1","H10","H100",..: 1 1 1 2 3 3 3 3 1114 1114 ...
# $ id_ind       : int  1 2 3 4 5 6 7 8 9 10 ...
# $ stratum      : Factor w/ 24 levels "1000","2000",..: 12 12 12 12 12 12 12 12 12 12 ...
# $ stratum_label: chr  "north_1_6" "north_1_6" "north_1_6" "north_1_6" ...
# $ sex          : int  1 2 1 2 1 1 2 2 1 1 ...
# $ cl_age       : Factor w/ 8 levels "(0,14]","(14,24]",..: 3 7 8 5 4 6 6 4 4 1 ...
# $ active       : num  1 1 0 1 1 1 1 1 1 0 ...
# $ income_hh    : num  30488 30488 30488 21756 29871 ...
# $ unemployed   : num  0 0 0 0 0 0 0 0 0 0 ...
# $ inactive     : num  0 0 1 0 0 0 0 0 0 1 ...
#-------------------------------------------------------------------------------
# Precision constraints
cv <- as.data.frame(list(DOM=c("DOM1","DOM2"),
                         CV1=c(0.02,0.03),
                         CV2=c(0.03,0.06),
                         CV3=c(0.03,0.06),
                         CV4=c(0.05,0.08)))
cv
#    DOM  CV1  CV2  CV3  CV4
# 1 DOM1 0.02 0.03 0.03 0.05
# 2 DOM2 0.03 0.06 0.06 0.08
#-------------------------------------------------------------------------------
# Preparation
samp_frame <- pop
samp_frame$one <- 1
id_PSU <- "municipality"  
id_SSU <- "id_ind"        
strata_var <- "stratum"   
target_vars <- c("income_hh","active","inactive","unemployed")   
deff_var <- "stratum"     
domain_var <- "region"  
delta =  1       # households = survey units
minimum <- 50    # minimum number of SSUs to be interviewed in each selected PSU
deff_sugg <- 1.5 # suggestion for the deff value
inp <- prepareInputToAllocation1(samp_frame,
                                 id_PSU,
                                 id_SSU,
                                 strata_var,
                                 target_vars,
                                 deff_var,
                                 domain_var,
                                 minimum,
                                 delta,
                                 deff_sugg)
#-------------------------------------------------------------------------------
# Optimal allocation
alloc <- beat.2st(stratif = inp$strata, 
                  errors = cv, 
                  des_file = inp$des_file, 
                  psu_file = inp$psu_file, 
                  rho = inp$rho, 
                  deft_start = NULL,
                  effst = inp$effst, 
                  minPSUstrat = 2,
                  minnumstrat = 50)
#   iterations PSU_SR PSU NSR PSU Total  SSU
# 1          0      0       0         0 7887
# 2          1     25     100       125 8342
# 3          2     28     100       128 8344
#-------------------------------------------------------------------------------
# First stage selection
sample_1st <- select_PSU(alloc, type="ALLOC", pps=TRUE, plot=TRUE)
sample_1st$PSU_stats
#    STRATUM PSU PSU_SR PSU_NSR  SSU SSU_SR SSU_NSR
# 1     1000   2      2       0  287    287       0
# 2     2000   7      1       6  380     52     328
# 3     3000   4      0       4  204      0     204
# 4     4000   2      0       2  100      0     100
# 5     5000   2      2       0  219    219       0
# 6     6000   2      0       2  100      0     100
# 7     7000   2      0       2  100      0     100
# 8     8000   2      0       2  100      0     100
# 9     9000   1      1       0  558    558       0
# 10   10000   6      6       0  588    588       0
# 11   11000  22      2      20 1198    100    1098
# 12   12000   8      0       8  410      0     410
# 13   13000   1      1       0  704    704       0
# 14   14000   4      4       0  577    577       0
# 15   15000  23      3      20 1281    161    1120
# 16   16000  16      0      16  830      0     830
# 17   17000   1      1       0  157    157       0
# 18   18000   3      1       2  166     50     116
# 19   19000   6      0       6  308      0     308
# 20   20000   4      0       4  200      0     200
# 21   21000   1      1       0  127    127       0
# 22   22000   3      3       0  151    151       0
# 23   23000   4      0       4  204      0     204
# 24   24000   2      0       2  100      0     100
# 25   Total 128     28     100 9049   3731    5318
#-------------------------------------------------------------------------------
# Second stage selection
sample_2st <- select_SSU(df=pop,
                   PSU_code="municipality",
                   SSU_code="id_ind",
                   PSU_sampled=sample_1st$sample_PSU)
head(sample_2st)
# municipality id_ind region province   id_hh stratum stratum_label sex  cl_age active income_hh
# 1          101   4359  north  north_1 H182836   12000     north_1_6   2 (24,34]      1  15404.61
# 2          101   4374  north  north_1 H182839   12000     north_1_6   2  (0,14]      0  11173.46
# 3          101   4425  north  north_1 H182857   12000     north_1_6   2 (64,74]      1  12103.86
# 4          101   4434  north  north_1 H182861   12000     north_1_6   2  (0,14]      0  19357.71
# 5          101   4449  north  north_1 H182865   12000     north_1_6   2 (54,64]      1  13949.31
# 6          101   4458  north  north_1 H182869   12000     north_1_6   1 (34,44]      1  25395.19
# unemployed inactive   Prob_1st   Prob_2st   Prob_tot   weight SR nSR stratum_2
# 1          0        0 0.09944943 0.05944056 0.00591133 169.1667  0   1    120003
# 2          0        1 0.09944943 0.05944056 0.00591133 169.1667  0   1    120003
# 3          0        0 0.09944943 0.05944056 0.00591133 169.1667  0   1    120003
# 4          0        1 0.09944943 0.05944056 0.00591133 169.1667  0   1    120003
# 5          0        0 0.09944943 0.05944056 0.00591133 169.1667  0   1    120003
# 6          0        0 0.09944943 0.05944056 0.00591133 169.1667  0   1    120003
```


