---
output: github_document
keep_md: TRUE
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

[![License](http://img.shields.io/badge/license-GPL%20%28%3E=%202%29-brightgreen.svg?style=flat)](http://www.gnu.org/licenses/gpl-3.0.html)
[![CRAN](http://www.r-pkg.org/badges/version/R2BEAT)](https://cran.r-project.org/package=R2BEAT)
[![Downloads](http://cranlogs.r-pkg.org/badges/R2BEAT?color=brightgreen)](http://www.r-pkg.org/pkg/R2BEAT)


# R2BEAT: Multistage Sampling Design and PSUs selection

Multivariate optimal allocation for different domains in one and two stages stratified sample design. R2BEAT extends the Neyman (1934) – Tschuprow (1923) allocation method to the case of several variables, adopting a generalization of the Bethel’s proposal (1989). R2BEAT develops this methodology but, moreover, it allows to determine the sample allocation in the multivariate and multi-domains case of estimates for two-stage stratified samples. It also allows to perform Primary Stage Units selection.

*R2BEAT* easily manages all the complexity due to the optimal sample allocation in two-stage sampling design and provides several outputs for evaluating the allocation.
Its name stands for "R 'to' Bethel Extended Allocation for Two-stage". 
It is an extension of another open-source software called *Mauss-R* (Multivariate Allocation of Units in Sampling Surveys), implemented by ISTAT  researchers (https://www.istat.it/en/methods-and-tools/methods-and-it-tools/design/design-tools/mauss-r). 
*Mauss-R* determines the optimal sample allocation in multivariate and multi-domains estimation, for one-stage stratified samples. 
It extends the Neyman (@neyman)  Tschuprow (@tschprow) allocation method to the case of several variables, adopting a generalization of the Bethel's proposal (@bethel1989). 
To complete the suite of tools developed by Istat in order to cover the stratified sample design, we cite *SamplingStrata* (https://CRAN.R-project.org/package=SamplingStrata), that allows to jointly optimize both the stratification of the sampling frame and the allocation, still in the multivariate multidomain case (only for one-stage designs), and *MultiWay.Sample.Allocation*, that allows to determine the optimal sample allocation for multi-way stratified sampling designs and incomplete stratified sampling designs (https://www.istat.it/en/methods-and-tools/methods-and-it-tools/design/design-tools/multiwaysampleallocation).

For a complete illustration of the methodology see the vignette "R2BEAT - Methodology and use". 

A complete example is illustrated in the vignette "Two-stage sampling design workflow". 


## Installation

You can install the released version of R2BEAT from [CRAN](https://CRAN.R-project.org) with:

``` {r, eval=FALSE, echo=TRUE}
install.packages("R2BEAT")
```

or the last version of R2BEAT from github with:
``` {r, eval=FALSE, echo=TRUE}
install.packages("devtools")
devtools::install_github("barcaroli/R2BEAT")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r, eval=FALSE, echo=TRUE}
library(R2BEAT)
# Load example data
data(beat.example)
head(strata)
#   STRATUM       N DOM1  DOM2         M1         M2 COST CENS        S1        S2
# 1       1 2302353 REG1 PROV1 0.02097409 0.01732549    1    0 0.1432975 0.1304811
# 2       2  179562 REG1 PROV2 0.02304681 0.01997698    1    0 0.1500522 0.1399210
# 3       3  371802 REG2 PROV3 0.02284502 0.01894097    1    0 0.1494093 0.1363166
# 4       4  592303 REG2 PROV3 0.02375591 0.01077033    1    0 0.1522878 0.1032198
# 5       5  221687 REG2 PROV3 0.03215470 0.01088398    1    0 0.1764108 0.1037570
# 6       6  440613 REG2 PROV3 0.04496313 0.01967690    1    0 0.2072232 0.1388874
head(errors)
#    DOM  CV1  CV2
# 1 DOM1 0.10 0.99
# 2 DOM2 0.99 0.99
head(design)
#   STRATUM STRAT_MOS DELTA MINIMUM
# 1       1   2302353     1      48
# 2       2    179562     1      48
# 3       3    371802     1      48
# 4       4    592303     1      48
# 5       5    221687     1      48
# 6       6    440613     1      48
head(PSU_strat)
#   STRATUM PSU_MOS PSU_ID
# 1       1    2591      1
# 2       1    3808      2
# 3       1     465      3
# 4       1    1778      4
# 5       1     713      5
# 6       1    6378      6
head(rho)
#   STRATUM RHO_AR1   RHO_NAR1 RHO_AR2     RHO_NAR2
# 1       1       1 0.07563272       1 0.0005484989
# 2       2       1 0.08327200       1 0.0157297282
# 3       3       1 0.03444641       1 0.0001944330
# 4       4       1 0.05433433       1 0.0021201253
# 5       5       1 0.03926152       1 0.0020177360
# 6       6       1 0.02452771       1 0.0293768074
# Allocate the sample
allocation <- beat.2st(stratif=strata, 
                            errors=errors,
                            des_file=design, 
                            psu_file=PSU_strat,
                            rho=rho)
#    iterations PSU_SR PSU NSR PSU Total   SSU
# 1           0      0       0         0  6858
# 2           1     16      79        95 17170
# 3           2     42     161       203 14736
# 4           3     33     153       186 15273
# 5           4     36     155       191 15147
# 6           5     33     156       189 15274
# 7           6     36     155       191 15146
# 8           7     33     156       189 15273
# 9           8     36     155       191 15147
# 10          9     33     156       189 15274
# 11         10     36     155       191 15146
# 12         11     33     156       189 15273
# 13         12     36     155       191 15147
# 14         13     33     156       189 15274
# 15         14     36     155       191 15146
# 16         15     33     156       189 15273
# 17         16     36     155       191 15147
# 18         17     33     156       189 15274
# 19         18     36     155       191 15146
# 20         19     33     156       189 15273
# 21         20     36     155       191 15147
str(allocation)
# Selection of PSUs
allocat <- allocation$alloc[-nrow(allocation$alloc),]
str(allocat)
allocat$STRATUM <- as.character(as.numeric(allocat$STRATUM)*100)
unique(allocat$STRATUM)
PSU_strat$STRATUM <- PSU_strat$STRATUM*100
unique(PSU_strat$STRATUM)
sampled_PSU <- StratSel(dataPop= PSU_strat,
                       idpsu= ~ PSU_ID,
                       dom= ~ STRATUM,
                       final_pop= ~ PSU_MOS,
                       size= ~ PSU_MOS,
                       PSUsamplestratum= 1,
                       min_sample= 12,
                       min_sample_index= FALSE,
                       dataAll=allocat,
                       domAll= ~ factor(STRATUM),
                       f_sample= ~ ALLOC,
                       planned_min_sample= NULL,
                       launch= F)
sampled_PSU[[2]]
#    Domain SRdom nSRdom SRdom+nSRdom SR_PSU_final_sample_unit NSR_PSU_final_sample_unit
# 1     100   126     53          179                     8240                       660
# 2     200    20     15           35                      305                        99
# 3     300     2      8           10                       28                        37
# 4     400     6     16           22                      199                       288
# 5     500     1      6            7                       68                       203
# 6     600     6      8           14                       11                        81
# 7     700     1      6            7                      939                       381
# 8     800     4      5            9                      140                       473
# 9     900     2      8           10                      423                       498
# 10   1000     3      8           11                      751                       186
# 11   1100     1      3            4                       51                        97
# 12   1200    10     23           33                       88                       192
# 13   1300     3     17           20                       37                        70
# 14   1400     1      7            8                      113                       104
# 15   1500    26     30           56                       26                        70
# 16   1600     7     39           46                       98                        59
# 17   1700    25     39           64                       42                        95
# 18  Total   244    291          535                    11559                      3593
# 19   Mean                                                680                       211
```

